---
title: "rvo_project"
author: "Jensen brock"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#set seed
set.seed(42)

# Libraries
library(tidyverse)
library(readxl)
library(broom)
library(scales)
library(modelsummary)
library(MatchIt)
library(class)
library(caTools)
library(AICcmodavg)


# Data
data_raw <- read_excel("data/Personal_Loans_Dataset_(Data).xlsx") 


data_clean <- data_raw %>% 
  mutate(
    possible_payout = case_when(
                    Lender == "A" ~ 250
                  , Lender == "B" ~ 350
                  , Lender == "C" ~ 150
                                      )
    , net_income = Monthly_Gross_Income - Monthly_Housing_Payment) %>% 
  rename(
      user_id = `User ID`
    , reason = Reason
    , loan_amount = Loan_Amount
    , fico_score = FICO_score
    , fico_score_group= Fico_Score_group
    , employment_status = Employment_Status
    , employment_sector = Employment_Sector
    , monthly_gross_inc = Monthly_Gross_Income
    , house_pmt_mo = Monthly_Housing_Payment
    , ever_bankrupt_foreclose = Ever_Bankrupt_or_Foreclose
    , lender = Lender
    , approved = Approved
  )

```


# Explore the Data
```{r summary}
summary(data_clean)
```

## Review breakdown of key numeric variables by primary dimensions
```{r}
data_clean %>%
  group_by(lender) %>% 
  summarise(
      approval_rate = mean(approved)
    , avg_loan_amount = mean(loan_amount)
    , avg_fico = mean(fico_score)
    , avg_inc = mean(monthly_gross_inc)
    , pct_bankrupt_foreclosed = mean(ever_bankrupt_foreclose)
    , payout = mean(possible_payout)
    , n_apps = n()
  )
```


```{r summary-fico-group}
data_clean %>%
  group_by(fico_score_group) %>% 
  summarise(
      avg_loan_amount = mean(loan_amount)
    , avg_fico = mean(fico_score)
    , avg_inc = mean(monthly_gross_inc)
    , approval_rate = mean(approved)
    , pct_bankrupt_foreclosed = mean(ever_bankrupt_foreclose)
  ) %>% 
  arrange(avg_fico)
```
```{r summary-employment-status}
data_clean %>%
  group_by(employment_status) %>% 
  summarise(
      avg_loan_amount = mean(loan_amount)
    , avg_fico = mean(fico_score)
    , avg_inc = mean(monthly_gross_inc)
    , approval_rate = mean(approved)
    , pct_bankrupt_foreclosed = mean(ever_bankrupt_foreclose)
  ) %>% 
  arrange(approval_rate)
```

```{r summary-emp-sector}
data_clean %>%
  group_by(employment_sector) %>% 
  summarise(
      avg_loan_amount = mean(loan_amount)
    , avg_fico = mean(fico_score)
    , avg_inc = mean(monthly_gross_inc)
    , approval_rate = mean(approved)
    , pct_bankrupt_foreclosed = mean(ever_bankrupt_foreclose)
  ) %>% 
  arrange(approval_rate)
```

```{r summary-loan-reason}
data_clean %>%
  group_by(reason) %>% 
  summarise(
      avg_loan_amount = mean(loan_amount)
    , avg_fico = mean(fico_score)
    , avg_inc = mean(monthly_gross_inc)
    , approval_rate = mean(approved)
    , pct_bankrupt_foreclosed = mean(ever_bankrupt_foreclose)
  ) %>% 
  arrange(approval_rate)

```
## Review Distributions
```{r fico-hist}
data_clean %>% 
  ggplot(mapping = aes(x = fico_score)) +
  geom_histogram(
              boundary = 600
            , bins = 12
            , color = "white")
```

```{r loan-amount-hist}
data_clean %>% 
  ggplot(mapping = aes(x = loan_amount)) +
  geom_histogram(
              binwidth = 10000
            , boundary = 0
            , color = "white") +
  scale_x_continuous(breaks = seq(0, max(data_clean$loan_amount)
                            , by = 10000)
                     , labels = scales::dollar)
```

```{r inc-hist}
data_clean %>% 
  ggplot(mapping = aes(x = monthly_gross_inc)) +
  geom_histogram(
              binwidth = 1500
            , boundary = 0
            , color = "white") +
  scale_x_continuous(breaks = seq(0, max(data_clean$monthly_gross_inc)
                            , by = 3000)
                     , labels = scales::dollar)
```

# Linear Probability model
This is a simple model that generates the probability of a loan application 
being approved. This is helpful for gaining understanding of how the various
factors influence approval, but this would not be a useful prediction model
due to the limitations of the Linear Probability Model.

Secondly we learned from the LPM that loan reason is not a significant predictor
of loan approval by itself in an LPM model. In the sections below I will use AIC
to find the best fit model.
```{r simple-lpm}
lpm <- lm(approved ~ lender + fico_score + net_income + loan_amount
                    + reason + employment_status, data = data_clean)
summary(lpm)
```

# Logit Regression Model

## Splitting Data into Training & Test
```{r split-data}
train <- data_clean %>% 
  slice_sample(prop = 0.7)

test <- data_clean %>% 
  anti_join(train)
  
```

## Train Logit Model
```{r logit-model-v1}
logit_model_1 <- glm(approved ~ lender + fico_score + net_income + loan_amount
                    + employment_status + reason
                , data = train
                , family = binomial(link = "logit"))

summary(logit_model_1)
```

```{r logit-model-v2}
logit_model_2 <- glm(approved ~ lender + fico_score + net_income + loan_amount
                    + employment_status + reason + reason*lender
                , data = train
                , family = binomial(link = "logit"))

summary(logit_model_2)
```

```{r logit-model-v3}
logit_model_3 <- glm(approved ~ lender + fico_score + net_income + loan_amount
                    + employment_status
                , data = train
                , family = binomial(link = "logit"))

summary(logit_model_3)
```

```{r logit-model-v4}
logit_model_4 <- glm(approved ~ lender + fico_score + net_income + loan_amount
                    + employment_status + lender*fico_score
                , data = train
                , family = binomial(link = "logit"))

summary(logit_model_4)
```

```{r logit-model-v5}
logit_model_5 <- glm(approved ~ lender + fico_score + net_income + loan_amount
                  + employment_status + lender*fico_score + lender*net_income
                , data = train
                , family = binomial(link = "logit"))

summary(logit_model_5)
```

```{r logit-model-final}
logit_model <- glm(approved ~ lender + fico_score + net_income + loan_amount
                  + employment_status + lender*fico_score + lender*net_income
                   + loan_amount*fico_score
                , data = train
                , family = binomial(link = "logit"))

summary(logit_model)
```

### Model fit test with AIC
the more simple model that omits reason is the best fit model.
```{r compare-models}
models <- list(logit_model_1, logit_model_2, logit_model_3
               , logit_model_4, logit_model_5, logit_model)
model_names <- c("Fist Model", "Second Model", "Third Model"
                 , "Fourth Model", "Fifth Model", "Final Model")

aictab(cand.set = models, modnames = model_names)
```

## Fit test data to model
```{r}
testing <- test
testing$prob <- predict.glm(logit_model, newdata = test, type = "response")


test %>% 
  mutate(output_test = if_else(prob>=0.5, 1, 0)
       , correct = if_else(approved == output_test, 1,0)) %>% 
  group_by(output_test) %>% 
  summarise(count = n(),
            correct = mean(correct))
```

## test splitting data by lender 
```{r split-data, message=FALSE}
data_a <- data_clean %>% 
  filter(lender == "A")

    train_a <- data_a %>% 
      slice_sample(prop = 0.7)
    
    test_a <- data_a %>% 
      anti_join(train_a)

data_b <- data_clean %>% 
  filter(lender == "B")

    train_b <- data_b %>% 
      slice_sample(prop = 0.7)
    
    test_b <- data_b %>% 
      anti_join(train_b)

data_c <- data_clean %>% 
  filter(lender == "C")

    train_c <- data_c %>% 
      slice_sample(prop = 0.7)
    
    test_c <- data_c %>% 
      anti_join(train_c)
```

### Logit Models
```{r}
logit_model_a <- glm(approved ~ fico_score + net_income + loan_amount
                  + employment_status + loan_amount*fico_score
                , data = train_a
                , family = binomial(link = "logit"))

logit_model_b <- glm(approved ~ fico_score + net_income + loan_amount
                  + employment_status + loan_amount*fico_score
                , data = train_b
                , family = binomial(link = "logit"))

logit_model_c <- glm(approved ~ fico_score + net_income + loan_amount
                  + employment_status + loan_amount*fico_score
                , data = train_c
                , family = binomial(link = "logit"))
```

```{r fit-a}
test_a$prob <- predict.glm(logit_model_a, newdata = test_a, type = "response")


test_a %>% 
  mutate(output_test = if_else(prob>=0.5, 1, 0)
       , correct = if_else(approved == output_test, 1,0)) %>% 
  group_by(output_test) %>% 
  summarise(count = n(),
            correct = mean(correct))
```

```{r fit-b}
test_b$prob <- predict.glm(logit_model_b, newdata = test_b, type = "response")


test_b %>% 
  mutate(output_test = if_else(prob>=0.5, 1, 0)
       , correct = if_else(approved == output_test, 1,0)) %>% 
  group_by(output_test) %>% 
  summarise(count = n(),
            correct = mean(correct))
```

```{r fit-c}
test_c$prob <- predict.glm(logit_model_c, newdata = test_c, type = "response")


test_c %>% 
  mutate(output_test = if_else(prob>=0.5, 1, 0)
       , correct = if_else(approved == output_test, 1,0)) %>% 
  group_by(output_test) %>% 
  summarise(count = n(),
            correct = mean(correct))
```

```{r}

```




