---
title: "rvo_project"
author: "Jensen brock"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#set seed
set.seed(42)

# Libraries
library(tidyverse)
library(kableExtra)
library(readxl)
library(broom)
library(scales)
library(MatchIt)
library(class)
library(caTools)
library(AICcmodavg)


# Data
data_raw <- read_excel("data/Personal_Loans_Dataset_(Data).xlsx")


data_clean <- data_raw %>% 
  mutate(
    possible_payout = case_when(
                    Lender == "A" ~ 250
                  , Lender == "B" ~ 350
                  , Lender == "C" ~ 150
                                      )
    , net_income = Monthly_Gross_Income - Monthly_Housing_Payment) %>% 
  rename(
      user_id = `User ID`
    , reason = Reason
    , loan_amount = Loan_Amount
    , fico_score = FICO_score
    , fico_score_group= Fico_Score_group
    , employment_status = Employment_Status
    , employment_sector = Employment_Sector
    , monthly_gross_inc = Monthly_Gross_Income
    , house_pmt_mo = Monthly_Housing_Payment
    , ever_bankrupt_foreclose = Ever_Bankrupt_or_Foreclose
    , lender = Lender
    , approved = Approved
  )

```


# I: Variable Overview
##Explore the Data
### Summary Tables
```{r general-summary}
summary_data <- data_clean %>% 
  summarise(
    approval_rate = scales::percent(mean(approved))
  , avg_fico_score = round(mean(fico_score),2)
  , avg_loan_amount = scales::dollar(mean(loan_amount))
  , avg_income = scales::dollar(mean(monthly_gross_inc))
  , avg_house_pmt = scales::dollar(mean(house_pmt_mo))
  , house_inc_ratio = scales::percent(sum(house_pmt_mo)/sum(monthly_gross_inc))
  )

summary_data
```
```{r summary-by-emp-status}
data_clean %>% 
  group_by(employment_status) %>% 
  summarise(
    approval_rate = scales::percent(mean(approved))
  , avg_fico_score = round(mean(fico_score),2)
  , avg_loan_amount = scales::dollar(mean(loan_amount))
  , avg_income = scales::dollar(mean(monthly_gross_inc))
  , avg_house_pmt = scales::dollar(mean(house_pmt_mo))
  )
```

```{r lender-breakdown}
lender_breakdown <- data_clean %>% 
  group_by(lender) %>% 
  summarise(n = n()
          , approval_rate = mean(approved)
          , avg_fico_score = mean(fico_score)
          , avg_loan_amount = mean(loan_amount))

lender_breakdown
```


```{r credit-score-breakdown, message=FALSE}
credit_score_summary <- data_clean %>% 
  group_by(lender, fico_score_group) %>% 
  summarise(n = n()
          , approval_rate = mean(approved)
          , avg_fico_score = mean(fico_score)
          , avg_loan_amount = mean(loan_amount)) %>% 
  arrange(factor(fico_score_group, levels = c("excellent", "very_good"
                                              , "good", "fair", "poor"))) 

credit_score_summary 
```

```{r fico-group-summary}
lender_fico_group <- data_clean %>% 
  mutate(fico_score_excellent = if_else(fico_score_group=="excellent", 1, 0)
       , fico_score_very_good = if_else(fico_score_group=="very_good", 1, 0)
       , fico_score_good = if_else(fico_score_group=="good", 1, 0)
       , fico_score_fair = if_else(fico_score_group=="fair", 1, 0)
       , fico_score_poor = if_else(fico_score_group=="poor", 1, 0)) %>% 
  group_by(lender) %>% 
  summarise(excellent_percent = mean(fico_score_excellent)
          , very_good_percent = mean(fico_score_very_good)
          , good_percent = mean(fico_score_good)
          , fair_percent = mean(fico_score_fair)
          , poor_percent = mean(fico_score_poor)
          , total_applications = n()
  )

lender_fico_group
```

### Data Distributions
```{r fico-hist}
data_clean %>% 
  ggplot(mapping = aes(x = fico_score)) +
  geom_histogram(
              boundary = 600
            , bins = 12
            , color = "#FFFFFF"
            , fill = "#300CE9") +
  labs(title = "FICO Score Distribution") +
  xlab("FICO Score") +
  ylab("Count")
```

#### Loan Amount
```{r loan-amount-hist}
data_clean %>% 
  ggplot(mapping = aes(x = loan_amount)) +
  geom_histogram(
              binwidth = 10000
            , boundary = 0
            , color = "#FFFFFF"
            , fill = "#ED6B3E") +
  scale_x_continuous(breaks = seq(0, max(data_clean$loan_amount)
                            , by = 10000)
                     , labels = scales::dollar) +
  labs(title = "Loan Amount Distribution") +
  xlab("Loan Amount") +
  ylab("Count")
```

#### Income
```{r inc-hist}
data_clean %>% 
  ggplot(mapping = aes(x = monthly_gross_inc)) +
  geom_histogram(
              binwidth = 1500
            , boundary = 0
            , color = "#FFFFFF"
            , fill = "#168039") +
  scale_x_continuous(breaks = seq(0, max(data_clean$monthly_gross_inc)
                            , by = 3000)
                     , labels = scales::dollar) +
  labs(title = "Income Distribution") +
  xlab("Monthly Gross Income") +
  ylab("Count")
```


## Predictive Modeling
Identify statistically significant variables to predict loan approval. Logistic
Regression is used to model the probability that a loan will be approved.

### Logistic Regression
```{r data-setup, message=FALSE}
# Training & Test Data
train <- data_clean %>% 
  slice_sample(prop = 0.7)

test <- data_clean %>% 
  anti_join(train)
  
# Model Training
logit_model_1 <- glm(approved ~ lender + fico_score + monthly_gross_inc 
                     + house_pmt_mo + loan_amount + employment_status + reason
                , data = train
                , family = binomial(link = "logit"))

logit_model_2 <- glm(approved ~ lender + fico_score + net_income 
                     + monthly_gross_inc + house_pmt_mo + loan_amount 
                     + employment_status + reason + reason*lender
                , data = train
                , family = binomial(link = "logit"))

logit_model_3 <- glm(approved ~ lender + fico_score + monthly_gross_inc 
                     + house_pmt_mo + loan_amount  + employment_status
                , data = train
                , family = binomial(link = "logit"))

logit_model_4 <- glm(approved ~ lender + fico_score + monthly_gross_inc 
                     + house_pmt_mo + loan_amount + employment_status 
                     + lender*fico_score
                , data = train
                , family = binomial(link = "logit"))

logit_model_5 <- glm(approved ~ lender + fico_score + monthly_gross_inc 
                     + house_pmt_mo + loan_amount + employment_status 
                     + lender*fico_score + lender*monthly_gross_inc
                , data = train
                , family = binomial(link = "logit"))

logit_model <- glm(approved ~ lender + fico_score + monthly_gross_inc 
                    + house_pmt_mo + loan_amount + employment_status 
                    + lender*fico_score + lender*monthly_gross_inc
                    + loan_amount*fico_score
                , data = train
                , family = binomial(link = "logit"))
```
### Summary Stats Final Model
```{r}
tidy(logit_model)
```


### Model fit test with AIC
```{r compare-models}
models <- list(logit_model_1, logit_model_2, logit_model_3
               , logit_model_4, logit_model_5, logit_model)
model_names <- c("Fist Model", "Second Model", "Third Model"
                 , "Fourth Model", "Fifth Model", "Final Model")

aictab(cand.set = models, modnames = model_names)
```

### Predict test data with model
```{r }
test$prob <- predict.glm(logit_model, newdata = test, type = "response")

```


```{r model-accuracy}
test %>% 
  mutate(is_approved = if_else(prob>0.5, 1, 0)
       , correct = if_else(approved == is_approved, 1,0)) %>% 
  group_by(is_approved) %>% 
  summarise(count = n(),
            pct_correct = mean(correct))
```

```{r lender-prob}
test %>% 
  group_by(lender) %>% 
  summarise(
        approval_probability = mean(prob)
     ,  actual_approval_rate = mean(approved)
            )
```


```{r distribution-lender-pred}
ggplot(test, mapping = aes(x = prob, color = lender)) +
  geom_density() +
  scale_x_continuous(name = "Probability of Loan Approval"
                     , labels = scales::percent)
```

```{r by-fico-group}
ggplot(test, mapping = aes(x = prob, color = fico_score_group)) +
  geom_density() +
  scale_x_continuous(name = "Probability of Loan Approval"
                     , labels = scales::percent)
```

```{r by-fico-group}
ggplot(test, mapping = aes(x = prob, color = fico_score_group)) +
  geom_density() +
  facet_wrap(vars(lender)) +
  scale_x_continuous(name = "Probability of Loan Approval"
                     , labels = scales::percent) +
  theme(legend.position = "bottom")
```

```{r}
ggplot(test, mapping = aes(x = loan_amount, y = prob, color = lender)) +
  geom_smooth(se = FALSE) +
  facet_wrap(vars(factor(fico_score_group, levels = c("excellent", "very_good"
                                              , "good", "fair", "poor")))) +
  scale_y_continuous(name = "Probability of Loan Approval"
                     , labels = scales::percent) +
  scale_x_continuous(name = "Loan Amount"
                   , labels = scales::dollar)
```











